Sexual selection improves population fitness: a meta-analysis

### 2018-09-26 Manuscripts/Manuscript_MetaAnalysis.Rmd
Luke changes to MS

---
title: "Sexual selection improves population fitness: a meta-analysis"
author: 
- #Justin G. Cally\textsuperscript{§*}, Devi Stuart-Fox\textsuperscript{§} and Luke Holman\textsuperscript{§}
- 
- #'\textsuperscript{*}\textit{justin.g.cally@gmail.com}'
output:
  pdf_document: default
  html_document: default
  word_document:
      pandoc_args: ["--smart"]
bibliography: Bibliography/Bibliography.bib
csl: Bibliography/NComms.csl
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
  \rhead[]{Sexual selection improves fitness} \usepackage{titlefoot}
  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\onehalfspacing
abstract: Sexual selection has manifold ecological and evolutionary consequences, making its net effect on population fitness difficult to predict theoretically. A powerful means to test whether sexual selection increases or decreases population fitness is to manipulate the opportunity for sexual selection in two or more populations, allow evolution to proceed, and then compare the average fitness between the treatment groups. Here, we synthesise 459 effect sizes from 65 experimental evolution studies that tested for effects of sexual selection on adaptation. We find that sexual selection tends to elevate the mean and reduce the variance for many fitness traits; especially for females evolving under stressful conditions. We conclude that the beneficial population-level consequences of sexual selection are typically stronger than the harmful ones, and that the effects of sexual selection on total selection, genetic variance, and/or evolvability can differ markedly between sexes and environments. We discuss the implications of these results for conservation and evolutionary biology.
---
<!-- \maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.} -->

@@

Sexual selection, defined as selection resulting from competition for mates or their gametes[@RN139], is a ubiquitous evolutionary force that has profoundly shaped the natural world. As far back as Darwin[@RN310], researchers have theorised that sexual selection can change the average absolute fitness of individuals in a population (henceforth termed ‘population fitness’[@RN269; @RN268]). However, opinion is divided over whether the net effect on population fitness is positive or negative[@RN42; @RN25; @RN116; @RN95]. _Prima facie_, one might predict that sexual selection would have no effect on population fitness, since it does not matter which individuals of the faster-reproducing sex (typically males) succeed in mating, so long as some do[@RN142]. However, when genotypes with high mating or fertilisation success also have superior breeding values for traits that affect population fitness (e.g. survival, parental care, female fecundity, or success in interspecific competition), sexual selection is predicted to elevate population fitness by causing a correlated response in these other traits[@RN95]. In essence, the demographically-limiting sex (typically females) benefits from a gene pool that has been purged of harmful alleles through sexual selection on the non-limiting sex (typically males), with the benefit to population fitness potentially large[@RN94; @RN198; @RN58].

Conversely, sexual selection can decrease population fitness if male sexually-selected traits are negatively genetically correlated with female fitness (intralocus sexual conflict)[@RN187; @RN146; @RN237; @RN147]. Additionally, sexual selection frequently favours phenotypes that reduce population fitness but benefit the individuals expressing them, such as harassment or infanticide by mate-seeking males[@RN6; @RN267; @RN37], as well as investment in costly sexual signals and weaponry at the expense of parental care (interlocus sexual conflict)[@RN314; @RN276; @RN315]. Given these conflicting theoretical expectations and empirical results, it remains unclear whether sexual selection tends to have a net benefit or cost to population fitness[@RN42; @RN25; @RN116; @RN95].

@@

Theoretical and empirical work has highlighted a number of factors that affect the relationship between sexual selection and population fitness. For example, the genetic correlation between female fecundity and male mating/fertilisation success varies in sign and magnitude between species[@RN298] and even between conspecific populations[@RN201], implying that sexual selection on males increases mean female fitness in some species and populations but not others. These inconsistencies could derive from differences in allele frequencies, or environmental differences that alter how genotype relates to fitness. Importantly, several studies have hypothesised that populations should display a more positive genetic correlation between male and female fitness – and thus potentially between mating/fertilisation success and population fitness – in novel or fluctuating environments, relative to stable environments[@RN187; @RN173; @RN74; @RN257]. These predictions arise because stable environments create consistent selection, preferentially eroding genetic variation at sexually-concordant loci (i.e. loci where the fittest genotype is the same in both sexes) and leaving behind variation at sexually antagonistic loci. Although there are several systematic reviews of the effects of sexual selection on offspring fitness (see ref.[@RN183]), none that we know of have addressed whether sexual selection is more beneficial to population fitness in stressful conditions or novel environments, despite the relevance of this question to conservation biology.

Here, we synthesise the empirical literature on sexual selection and population fitness using formal meta-analysis. We focus exclusively on experimental evolution studies that manipulated the presence or strength of sexual selection, and then measured some fitness component, since these provide a particularly strong test of the hypothesis that sexual selection affects the average fitness of populations. Our principle aims were to measure the average net effect of sexual selection on the population-level mean and variance for various fitness components, and to test whether this effect varies between stressful and benign environments.

@@

We retrieved 459 effect sizes from 65 studies. 94 effect sizes were collected from populations evolving under stressful conditions, while 335 were measured for those evolving in benign conditions. 189 of the effect sizes came from measurements made on males, 219 on females, and the remaining 51 from measurements of a mixed-sex sample of individuals. Most effect sizes in our dataset came from studies that manipulated sexual selection by completely removing it in one treatment via enforced random monogamy (*n* = 241); other effect sizes (*n* = 218) derived from alternative manipulations, such as changing the adult sex ratio. In total, we obtained effect sizes for 20 different fitness traits, with reproductive success (*n* = 156) and offspring viability (*n* = 56) being the most commonly-measured traits. We classified 216 effect sizes as direct measures of fitness, 147 as indirect, and the remaining 96 effect sizes as ambiguous (see Methods, Fig. 1 and Table S1). Specifically, direct measures were components of reproduction or long-term viability (*e.g.*, reproductive success, offspring viability and extinction rate), indirect measures were traits often used as a proxy of fitness but that do not directly measure aspects of success in reproduction or population viability (*e.g.*, lifespan, mating success and ejaculate quality/production); and ambiguous measures were those reported to have an unclear or variable association with fitness (*e.g.*, body size, mating duration and early fecundity). Tables S4-S5 give a detailed description of our dataset.

@@

The grand mean for all effect sizes was positive (REML $\beta$ = 0.23, 95 % CIs: 0.01 to 0.44, *p* = 0.038; Bayesian $\beta$ = 0.23, 95 % CIs: -0.06 to 0.51, $BF_{>0}$ = 18), indicating that sexual selection typically had a net positive effect on the majority of populations and fitness components so far studied. Moreover, the effect sizes associated with the manipulation of sexual selection varied between different fitness traits. Sexual selection had a beneficial effect on most fitness traits, including all of the fitness traits that we classified as being ‘direct’ measurements of population fitness: extinction rate (REML $\beta$ = 0.37, 95 % CIs: 0.01 to 0.73; Bayesian $\beta$ = 0.37, 95 % CIs: -0.02 to 0.79, $BF_{>0}$ = 30.13; *n* = 4), offspring viability (REML $\beta$ = 0.20, 95 % CIs: 0.042 to 0.36; Bayesian $\beta$ = 0.21, 95 % CIs: -0.04 to 0.47, $BF_{>0}$ = 24.89; *n* = 56) and reproductive success (REML $\beta$ = 0.17, 95 % CIs: 0.013 to 0.32; Bayesian $\beta$ = 0.18, 95 % CIs: -0.06 to 0.44, $BF_{>0}$ = 15.53; *n* = 156; Fig. 1; Table S6). Sexual selection significantly reduced two fitness components, namely immunity (REML $\beta$ = -0.49, 95 % CIs: -0.72 to -0.26; Bayesian $\beta$ = -0.49, 95 % CIs: -0.78 to -0.19, $BF_{>0}$ = 0.0054; *n* = 35) and body condition (REML $\beta$ = -1.3, 95 % CIs: -1.9 to -0.67; Bayesian $\beta$ = -1.3, 95 % CIs: -1.9 to -0.63, $BF_{>0}$ = 0.00025, *n* = 1). Fig. 1 presents Bayesian estimates for each fitness component and Table S6 contains both Bayesian and REML predictions with sample sizes and BF’s.

@@

The effect size estimates we recovered were highly heterogeneous ($I^2$ = 94.9 %, 95 % CIs: 93 to 96%), reflecting the large differences in experimental procedures, study species, and fitness components included in our meta-analysis [@RN208]. Most of the observed heterogeneity stemmed from between-study differences ($I^2_{study}$ = 61.5 %, 95 % CIs: 42.9 % to 78.9 %), with variation between fitness components accounting for comparatively less heterogeneity ($I^2_{fitness\:components}$ = 33.5 %, 95 % CIs: 15.8 % to 52.7 %). The effect of taxon varied between REML and Bayesian models: the REML model suggested no effect ($I^2_{taxon}$ = 0), while the Bayesian model implied a non-zero effect ($\sigma^2$ = 0.145 [0.006-0.517]; Table S10). Variation among taxa is explored further in the Supplementary Material, (Figure S3 and Table S8).

@@

By applying meta-analysis to log coefficient of variation ratios (*lnCVRs*)[@RN177], we found that sexual selection reduces phenotypic variation under certain conditions (Fig. 2b). Specifically, phenotypic variance was significantly reduced by sexual selection for fitness components measured in females under stressful conditions (*lnCVR* = -1.0, 95% CIs: -1.4 to -0.69). By contrast, we found no significant effect of sexual selection on phenotypic variance in males, or for either sex under benign conditions (Fig. 2b; Tables S13-S16). However, as expected from the results in females, there was a non-significant trend for a reduction in phenotypic variance in mixed-sex samples measured under stressful conditions (*lnCVR* = -0.44, 95% CIs: -0.80 to -0.081; Fig. 2b).

As in the meta-analysis of trait means, there was high heterogeneity in the estimates of *lnCVR* ($I^2$ = 95.4 %, 95 % CIs: 93 % to 97.1 %). Heterogeneity in the dataset was due to variability between studies ($I^2_{study}$ = 40.1 %, 95 % CIs: 23.5 % to 60.2 %), fitness components ($I^2_{fitness\:components}$ = 38.7 %, 95 % CIs: 17.1 % to 60.7 %) and taxon ($I^2_{taxon}$ = 16.6 %, 95 % CIs: 3.1 % to 37.2 %).
<!-- We don't use Bayesian predictions for lnCVR, which is inconsistent with previous results -->

@@

The funnel plot of effect sizes was asymmetrical, suggesting that some publication bias might be present (Fig. 3a; Egger’s test: *z* = 6.2, *p* < 0.0001). Specifically, there was a moderate excess of low-powered studies in which sexual selection had a more positive effect on the fitness component than average, implying that low-powered studies are more likely to be published if they report statistically significant fitness benefits of sexual selection. There was no significant relationship between effect size and journal impact factor (Fig. 3b; $t_{437}$ = 1.2, *p* = 0.25) or year of publication (Fig. 3c; $t_{437}$ = -1.0, *p* = 0.31); thus, we found no evidence that effect size dictates the likelihood of publication in high-profile journals, or that effect sizes have diminished as the field has matured[@RN216; @RN217].

```{r, echo = FALSE, fig.cap="Forest plot showing the effect of manipulation of sexual selection, for 20 fitness components from 65 studies. Effect sizes (*n* = 459) varied between fitness components, and were more often positive than negative. Predicted average values are presented alongside each fitness component; those in bold have 95 % confidence intervals that do not cross zero. The estimates presented here are Bayesian predictions, which appear slightly more conservative than REML predictions (Table S6)."}
knitr::include_graphics("Figures/ForestPlot_edit3.pdf") #Not relative path
```

@@

```{r, echo = FALSE, fig.cap= "Tests of publication bias are mixed and suggest publication bias may be present. (**a**) Inspection and statistical tests of the funnel plot reveal large amounts of heterogeneity in the dataset with asymmetry from increased low-powered, large effect studies. (**b**) No significant correlation exists between journal impact factor and effect size, (**c**) additionally there is no significant correlation between effect size and year of publication when testing for the time-lag bias. For (**b**) and (**c**) circle size is proportional to the weight (%) of the study, which is inversely proportional to the variance of the effect size."}
knitr::include_graphics("Figures/publicationbias.pdf") #Not relative path
```

@@

Our meta-analysis revealed that populations evolving under sexual selection often have higher values for multiple fitness traits, relative to populations where sexual selection was experimentally removed or weakened. Notably, sexual selection had small but beneficial effects on direct measures of fitness such as extinction rate, reproductive success (defined as the number of offspring produced) and the proportion of viable offspring. Fitness traits related to immunocompetence were an exception: sexual selection typically resulted in weaker immunity. The negative effect on immunity is interesting in light of the hypothesised trade-off between sexually-selected phenotypes and immunity, for example due to immunosuppressive effects of sex hormones[@RN323; @RN324]. Furthermore, the overall benefit of sexual selection was greater for females than males and this sex difference was magnified in stressful environments. Consistent with stronger selection on female fitness under stress, female phenotypic variance was substantially reduced under these conditions. These results suggest that sexual selection may be a key factor in population persistence under stressful conditions, such as fluctuating environmental change[@RN173] or spatial variability[@RN189]. Under such conditions, female reproductive output often acts as the limiting factor in staving off extinction[@RN188].

The results of the meta-analysis support predictions that sexual selection can improve population fitness and hence aid adaptation[@RN94; @RN173; @RN54; @RN141; @RN230; @RN198; @RN58]. Specifically, our results support arguments that positive genetic correlations between male mating success and female non-sexual traits are common, allowing females to benefit from a genome that has been ‘purged’ of deleterious alleles through intense selection on males[@RN94; @RN198; @RN95]. What is less clear is why sexual selection had a larger effect on female trait means and variances as opposed to males. Below we discuss possible explanations for this result, in light of the core principle that the speed and magnitude of adaptation depends on additive genetic variance, the structure of genetic covariance, and the strength of selection[@RN317; @RN318].

Firstly, the response of male fitness to sexual selection may be constrained in species where the heritability of male fitness is influenced by the sex determination system. Males are heterogametic in many of the species in our sample (e.g. in species with XY or XO sex determination systems); this can reduce father-to-son heritability relative to mother-to-daughter, since sons never inherit the larger sex chromosome from their fathers[@RN340; @RN262; @RN147; @RN263]. This imbalance of male phenotypes that depend on X-linked loci should slow adaptation relative to rapidly adapting female phenotypes[@RN263; @RN259]. However,  X-linkage is also likely to inflate the contribution of X-linked genes to fitness variance in the heterogametic sex. Under this view, X-linkage facilitates adaptation by increasing the efficacy of selection against recessive deleterious alleles and selection for rare (and recessive) beneficial alleles [@RN341; @RN339]. Given these countering viewpoints, there is no clear consensus on sex-based differences in heritability. For instance, a systematic review found no difference in mean heritability between male and female traits, however there was a male-biased skew in heritability of reproductive traits, which are more likely to be tied to fitness[@RN342]. Conversely, a recent GWAS in humans found a number of traits with higher heritability in females than males[@RN325].

Secondly, selection on males might be weaker than selection on females, resulting in slower adaptation following experimental manipulation of sexual selection. This explanation may initially seem implausible, because net selection on males is generally thought to be stronger than on females due to the cumulative effects of sexual and natural selection[@RN304; @RN303; @RN337; @RN338]. However, an oft-overlooked aspect is that selection might frequently be ‘softer’ on males and ‘harder’ on females[@RN319], because the local competitive environment is usually more important for males than it is for females. For instance, a mediocre male genotype can have high fitness provided it outcompetes its local rivals, while low-fitness female genotypes are likely to produce few offspring even when competing with other low-fitness females. Therefore, improvements in genetic quality should have stronger diminishing returns in males, possibly contributing to our finding that the genetic consequences of sexual selection lead to greater fitness benefits to females. Though this argument is speculative, we note that many experimental evolution designs exaggerate the sex difference in the nature of selection, relative to expectations for large, natural populations[@RN322; @RN277; @RN112]. For example, many studies[@RN110; @RN326; @RN300; @RN111; @RN227] have evolved insects in small sub-populations, each containing one female and multiple males, whose progeny are then randomly mixed to create the next generation; this ensures that successful males simply needed to outcompete their rival(s) in the same sub-population, while successful females needed to outcompete all other females in the entire population.

Our results suggest that the greater benefit of sexual selection to females than males is magnified in stressful environments. Recent work has emphasised that environmental stress should reduce the strength of sexually antagonistic selection relative to selection that is concordant between sexes. Theoretical models reaching this conclusion[@RN67; @RN173] have been supported by empirical studies[@RN187; @RN74; @RN243]; for example, one study found that high fitness males produced low fitness daughters under benign conditions but high fitness daughters under stress[@RN74]. However, other quantitative genetic studies have shown that stressful conditions do not always reduce sexual antagonism[@RN297; @RN200; @RN257]. Variation in effects of sexual selection in stressful environments may be due to potentially variable responses amongst taxa[@RN298] and environments. Notably, Connallon and Hall (2016)[@RN173] predict that the dynamics of environmental change alter the strength of sexual antagonism; for instance, gradual directional selection may facilitate indefinite sexual antagonism, while rapid cyclical change can swiftly remove it. Our meta-analysis suggests that under directional selection generally imposed by environmental stress, sexual antagonism is likely dampened; allowing sexual selection to facilitate adaptation and persistence.

>>>>

---
title: "Sexual selection improves population fitness: a meta-analysis"
author: 
- #Justin G. Cally\textsuperscript{§*}, Devi Stuart-Fox\textsuperscript{§} and Luke Holman\textsuperscript{§}
- 
- #'\textsuperscript{*}\textit{justin.g.cally@gmail.com}'
output:
  pdf_document: default
  html_document: default
  word_document:
      pandoc_args: ["--smart"]
bibliography: Bibliography/Bibliography.bib
csl: Bibliography/NComms.csl
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
  \rhead[]{Sexual selection improves fitness} \usepackage{titlefoot}
  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\onehalfspacing
abstract: Sexual selection has manifold ecological and evolutionary consequences, making its net effect on population fitness difficult to predict from first principles. Instead, one can empirically determine whether sexual selection increases or decreases population fitness by  manipulating the opportunity for sexual selection in replicate populations, allow evolution to proceed, and then comparing the evolved average fitness between the treatment groups. Here, we synthesise 459 effect sizes from 65 pertinent experimental evolution studies using meta-analysis. We find that sexual selection tends to elevate the mean and reduce the variance for many fitness traits, especially in females, and for populations evolving under stressful conditions. We conclude that the beneficial population-level consequences of sexual selection typically outweigh the harmful ones, and that the effects of sexual selection on total selection, genetic variance, and/or evolvability can differ markedly between sexes and environments. We discuss the implications of these results for conservation and evolutionary biology. \vspace{5mm} \par\noindent \textbf{Keywords:} Adaptation, Gender load, Monogamy, Mutation load, Sexual conflict. \vspace{5mm} \par\noindent \textbf{Word count:} XXXX in main text.
---
<!-- \maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.} -->

@@

Sexual selection, defined as selection resulting from competition for mates or their gametes[@RN139], is a ubiquitous evolutionary force that has profoundly shaped the natural world. As far back as Darwin[@RN310], researchers have theorised that sexual selection can change the average absolute fitness of individuals in a population (henceforth termed ‘population fitness’[@RN269; @RN268]). However, opinion is divided over whether the net effect on population fitness is positive or negative[@RN42; @RN25; @RN116; @RN95]. _Prima facie_, one might predict that sexual selection would have no effect on population fitness, since it does not matter which individuals of the faster-reproducing sex (typically males) succeed in mating, so long as some do[@RN142]. However, when genotypes with high mating or fertilisation success also have superior breeding values for traits that affect population fitness (e.g. survival, parental care, female fecundity, or success in interspecific competition), sexual selection is predicted to elevate population fitness by causing a correlated response in these other traits[@RN95]. In essence, the demographically-limiting sex (typically females) benefits from a gene pool that has been purged of harmful alleles through sexual selection on the non-limiting sex (typically males). Theoretically, the benefit to population fitness might could be large[@RN94; @RN198; @RN58].

Conversely, sexual selection can decrease population fitness if male sexually-selected traits are negatively genetically correlated with female fitness, producing intralocus sexual conflict[@RN187; @RN146; @RN237; @RN147]. Additionally, sexual selection frequently favours phenotypes that reduce population fitness but benefit the individuals expressing them, such as harassment or infanticide by mate-seeking males[@RN6; @RN267; @RN37], as well as investment in costly sexual signals and weaponry at the expense of parental care (\textit{inter}locus sexual conflict)[@RN314; @RN276; @RN315]. Given these conflicting theoretical expectations and empirical results, it remains unclear whether sexual selection tends to have a net benefit or cost to population fitness[@RN42; @RN25; @RN116; @RN95].

@@

A number of factors might influence the strength and sign of the correlation between sexual selection and population fitness. First, the genetic correlation between female fecundity and male mating/fertilisation success varies in sign and magnitude between species[@RN298] and even between conspecific populations[@RN201], implying that sexual selection on males increases mean female fitness in some species and populations but not others. These inconsistencies could derive from differences in allele frequencies, or environmental differences that alter how genotype relates to phenotype and fitness. Second, it has been hypothesised that populations should display a more positive genetic correlation between male and female fitness – and thus potentially between mating/fertilisation success and population fitness – in novel or fluctuating environments, relative to stable environments[@RN187; @RN173; @RN74; @RN257]. This is because stable environments create consistent selection, preferentially eroding genetic variation at sexually-concordant loci (i.e. loci where the fittest genotype is the same in both sexes) and leaving behind variation at sexually antagonistic loci. We know of no systematic reviews testing this latter theory, though it is has motivated several recent empirical tests (e.g.[@RN187; @RN200; @RN254]), and is relevant to conservation genetics.

Here, we synthesise the empirical literature on sexual selection and population fitness using formal meta-analysis. We focus exclusively on experimental evolution studies that manipulated the presence or strength of sexual selection, and then measured some fitness component, since these provide a particularly strong test of the hypothesis that sexual selection affects the average fitness of populations. Our principle aims were to measure the average net effect of sexual selection on the population-level mean and variance for various fitness components, to test whether this effect varies between stressful and benign environments, and identify key moderators of the effect size of sexual selection treatment.

@@

We retrieved 459 effect sizes from 65 studies. 94 effect sizes were collected from populations evolving under stressful conditions, while 335 were measured for those evolving in benign conditions. 189 of the effect sizes came from measurements made on males, 219 on females, and the remaining 51 from measurements of a mixed-sex sample of individuals. Most effect sizes in our dataset came from studies that manipulated sexual selection by completely removing it in one treatment via enforced random monogamy ($n$ = 241); other effect sizes ($n$ = 218) derived from alternative manipulations, such as changing the adult sex ratio. In total, we obtained effect sizes for 20 different fitness traits, with reproductive success ($n$ = 156) and offspring viability ($n$ = 56) being the most commonly-measured traits. We classified 216 effect sizes as direct measures of fitness, 147 as indirect, and the remaining 96 effect sizes as ambiguous (see Methods, Fig. 1 and Table S1). Specifically, direct measures were components of reproduction or long-term viability (*e.g.*, reproductive success, offspring viability and extinction rate), indirect measures were traits often used as a proxy of fitness but that do not directly measure aspects of success in reproduction or population viability (*e.g.*, lifespan, mating success and ejaculate quality/production); and ambiguous measures were those reported to have an unclear or variable association with fitness (*e.g.*, body size, mating duration and early fecundity). Tables S4-S5 give a detailed description of our dataset.

@@

The grand mean for all effect sizes was positive (REML $\beta$ = 0.23, 95% CIs: 0.01 to 0.44, $p$ = 0.038; Bayesian $\beta$ = 0.23, 95% CIs: -0.06 to 0.51, $BF_{>0}$ = 18), indicating that sexual selection typically had a net positive effect on the majority of populations and fitness components so far studied. Moreover, the effect sizes associated with the manipulation of sexual selection varied between different fitness traits. Sexual selection had a beneficial effect on most fitness traits, including all of the fitness traits that we classified as being ‘direct’ measurements of population fitness: extinction rate (REML $\beta$ = 0.37, 95% CIs: 0.01 to 0.73; Bayesian $\beta$ = 0.37, 95% CIs: -0.02 to 0.79, $BF_{>0}$ = 30.13; $n$ = 4), offspring viability (REML $\beta$ = 0.20, 95% CIs: 0.042 to 0.36; Bayesian $\beta$ = 0.21, 95% CIs: -0.04 to 0.47, $BF_{>0}$ = 24.89; $n$ = 56) and reproductive success (REML $\beta$ = 0.17, 95% CIs: 0.013 to 0.32; Bayesian $\beta$ = 0.18, 95% CIs: -0.06 to 0.44, $BF_{>0}$ = 15.53; $n$ = 156; Fig. 1; Table S6). Sexual selection significantly reduced two fitness components, namely immunity (REML $\beta$ = -0.49, 95% CIs: -0.72 to -0.26; Bayesian $\beta$ = -0.49, 95% CIs: -0.78 to -0.19, $BF_{>0}$ = 0.0054; $n$ = 35) and body condition (REML $\beta$ = -1.3, 95% CIs: -1.9 to -0.67; Bayesian $\beta$ = -1.3, 95% CIs: -1.9 to -0.63, $BF_{>0}$ = 0.00025, $n$ = 1). Fig. 1 presents Bayesian estimates for each fitness component and Table S6 contains both Bayesian and REML estimates, with sample sizes and test statistics.

@@

The effect size estimates we recovered were highly heterogeneous ($I^2$ = 94.9%, 95% CIs: 93-96%), reflecting the large differences in experimental procedures, study species, and fitness components included in our meta-analysis [@RN208]. Most of the observed heterogeneity stemmed from between-study differences ($I^2_{study}$ = 61.5%, 95% CIs: 42.9-78.9%), with variation between fitness components accounting for comparatively less heterogeneity ($I^2_{fitness\:components}$ = 33.5%, 95% CIs: 15.8-52.7%). The effect of taxon varied between REML and Bayesian models: the REML model suggested no effect ($I^2_{taxon}$ = 0), while the Bayesian model implied a non-zero effect ($\sigma^2$ = 0.145 [0.006-0.517]; Table S10). Variation among taxa is explored further in the Supplementary Material, (Figure S3 and Table S8).

@@

By applying meta-analysis to log coefficient of variation ratios (*lnCVRs*)[@RN177], we found evidence that sexual selection reduces phenotypic variation under certain conditions (Fig. 2b). Specifically, phenotypic variance was significantly reduced by sexual selection for fitness components measured in females under stressful conditions (*lnCVR* = -1.0, 95% CIs: -1.4 to -0.69). By contrast, we found no significant effect of sexual selection on phenotypic variance in males, or for either sex under benign conditions (Fig. 2b; Tables S13-S16). However, as expected from the results in females, there was a non-significant trend for a reduction in phenotypic variance in mixed-sex samples measured under stressful conditions (*lnCVR* = -0.44, 95% CIs: -0.80 to -0.081; Fig. 2b).

As in the meta-analysis of trait means, there was high heterogeneity in the estimates of *lnCVR* ($I^2$ = 95.4%, 95% CIs: 93-97%). Heterogeneity in the dataset was due to variability between studies ($I^2_{study}$ = 40.1 %, 95% CIs: 23.5 % to 60.2 %), fitness components ($I^2_{fitness\:components}$ = 38.7%, 95% CIs: 17.1-60.7%) and taxon ($I^2_{taxon}$ = 16.6, 95% CIs: 3.1-37.2%).
<!-- We don't use Bayesian predictions for lnCVR, which is inconsistent with previous results -->

@@

The funnel plot of effect sizes was asymmetrical, suggesting that some publication bias might be present (Fig. 3a; Egger’s test: *z* = 6.2, $p$ < 0.0001). Specifically, there was a moderate excess of low-powered studies in which sexual selection had a more positive effect on the fitness component than average, implying that low-powered studies are more likely to be published if they report statistically significant fitness benefits of sexual selection. There was no significant relationship between effect size and journal impact factor (Fig. 3b; $t_{437}$ = 1.2, $p$ = 0.25) or year of publication (Fig. 3c; $t_{437}$ = -1.0, $p$ = 0.31); thus, we found no evidence that effect size dictates the likelihood of publication in high-profile journals, or that effect sizes have diminished as the field has matured[@RN216; @RN217].

```{r, echo = FALSE, fig.cap="Forest plot showing the effect of manipulation of sexual selection, for 20 fitness components from 65 studies. Effect sizes ($n$ = 459) varied between fitness components, and were more often positive than negative. Predicted average values are presented alongside each fitness component; those in bold have 95% confidence intervals that do not cross zero. The estimates presented here are Bayesian predictions, which were slightly more conservative than the REML predictions (Table S6)."}
knitr::include_graphics("Figures/ForestPlot_edit3.pdf") #Not relative path
```

@@

```{r, echo = FALSE, fig.cap= "Tests of publication bias are mixed and suggest publication bias may be present. (**a**) Inspection and statistical tests of the funnel plot reveal large amounts of heterogeneity in the dataset with asymmetry from increased low-powered, large effect studies. (**b**) No significant correlation exists between journal impact factor and effect size, (**c**) additionally there is no significant correlation between effect size and year of publication when testing for the time-lag bias. For (**b**) and (**c**) point size is proportional to the precision of the effect size (i.e. the inverse of its variance)."} 
knitr::include_graphics("Figures/publicationbias.pdf") #Not relative path
```

@@

Our meta-analysis revealed that populations evolving under sexual selection often have higher values for multiple fitness traits, relative to populations where sexual selection was experimentally removed or weakened. Notably, sexual selection had small but beneficial effects on direct measures of fitness such as extinction rate, reproductive success (defined as the number of offspring produced) and the proportion of viable offspring. Fitness traits related to immunocompetence were an exception: sexual selection typically resulted in weaker immunity. This result is interesting in light of the hypothesised trade-off between sexually-selected phenotypes and immunity, for example due to immunosuppressive effects of sex hormones[@RN323; @RN324]. Furthermore, the overall benefit of sexual selection was greater for females than males, and this sex difference was magnified in stressful environments. Consistent with stronger selection on female fitness under stress, female phenotypic variance was substantially reduced when sexual selection was applied under stressful as opposed to benign conditions. These results suggest that sexual selection may contribute to population persistence under stressful conditions, such as fluctuating environmental change[@RN173] or spatial variability[@RN189], particularly since female reproductive output is often a limiting factor in population growth[@RN188].

The results of the meta-analysis support predictions that sexual selection can improve population fitness and accelerate adaptation[@RN94; @RN173; @RN54; @RN141; @RN230; @RN198; @RN58]. Specifically, our results support arguments that positive genetic correlations between male mating success and female non-sexual traits are common, allowing females to benefit from a genome that has been 'purged' of deleterious alleles through selection on males[@RN94; @RN198; @RN95]. What is less clear is why sexual selection had a larger effect on female trait means and variances as opposed to males (the sex in which sexual selection is typically strongest). Below we discuss possible explanations for this result, in light of the core principle that the speed and magnitude of adaptation depends on additive genetic variance, the structure of genetic covariance, and the strength of selection[@RN317; @RN318].

Firstly, the response of male fitness to sexual selection may be constrained in species where the heritability of male fitness is influenced by the sex determination system. Males are heterogametic in many of the species in our sample (e.g. in species with XY or XO sex determination systems); this can reduce father-to-son heritability relative to mother-to-daughter, since sons never inherit the larger sex chromosome from their fathers[@RN340; @RN262; @RN147; @RN263], which might slow adaptation in male relative to female traits[@RN263; @RN259]. However,  X-linkage is also likely to inflate the contribution of X-linked genes to fitness variance in the heterogametic sex. Under this view, X-linkage facilitates adaptation by increasing the efficacy of selection against recessive deleterious alleles and selection for rare (and recessive) beneficial alleles [@RN341; @RN339]. Perhaps as a result of these conflicting predictions, there seems to be no clear consensus regarding sex-based differences in heritability. For instance, a systematic review found no difference in mean heritability between male and female traits, although there was a male-biased skew in heritability of reproductive traits[@RN342]. Conversely, a recent association study in humans found a number of traits showing higher heritability in females than males[@RN325].

Secondly, selection on males might be weaker than selection on females, resulting in slower adaptation following experimental manipulation of sexual selection. This explanation may initially seem implausible, because net selection on males is generally assumed to be stronger than on females, due to the (actual or hypothesised) sex difference in sexual selection strength and inter-individual variance in fitness[@RN304; @RN303; @RN337; @RN338]. However, an oft-overlooked aspect is that selection might frequently be ‘softer’ on males and ‘harder’ on females[@RN319], because the local competitive environment is usually more important for males than it is for females. For instance, a mediocre male genotype can have high fitness provided it outcompetes its local rivals, while low-fitness female genotypes are likely to produce few offspring even when competing with other low-fitness females. Therefore, improvements in genetic quality should have stronger diminishing returns in males, possibly contributing to our finding that the genetic consequences of sexual selection lead to greater fitness benefits to females. Though this argument is speculative, we note that many experimental evolution designs exaggerate the sex difference in the nature of selection, relative to expectations for large, natural populations[@RN322; @RN277; @RN112]. For example, many studies[@RN110; @RN326; @RN300; @RN111; @RN227] have evolved insects in small sub-populations, each containing one female and multiple males, whose progeny are then mixed and randomly sampled to create the next generation; this design ensures that successful males simply needed to outcompete their rival(s) in the same sub-population, while each female's reproductive output is measured against the entire female population.

Our results suggest that the greater benefit of sexual selection to females than males is magnified in stressful environments. Recent work has emphasised that environmental stress should reduce the strength of sexually antagonistic selection relative to selection that is concordant between sexes. Theoretical models reaching this conclusion[@RN67; @RN173] have been supported by empirical studies[@RN187; @RN74; @RN243]; for example, one study found that high fitness males produced low fitness daughters under benign conditions but high fitness daughters under stress[@RN74]. However, other quantitative genetic studies have shown that stressful conditions do not always reduce sexual antagonism[@RN297; @RN200; @RN257]. Variation in effects of sexual selection in stressful environments may be due to potentially variable responses amongst taxa[@RN298] and environments. Notably, Connallon and Hall[@RN173] predict that the dynamics of environmental change alter the strength of sexual antagonism; for instance, gradual directional selection may facilitate indefinite sexual antagonism, while rapid cyclical change can swiftly remove it. Our meta-analysis suggests that under directional selection imposed by environmental stress, sexual antagonism is likely dampened; allowing sexual selection to facilitate adaptation and persistence.
